Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-19 main.tex
comment
sequentialisation

%All statements in this section are proved by rule permutation arguments.
\begin{thm}\label{thm:LK1-decompose}
  Let $\Gamma$ be a sequent. If\/ $\sqn\Gamma$ is provable in $\FOLK$ (as depicted on the left below) then there is a sequent
  $\Gamma'$ such that there is a derivation as shown on the right below:\vadjust{\vskip-2ex}
  \begin{equation*}
    \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\Gamma}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}
    \qquad
    \Longrightarrow
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
        \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}}
  \end{equation*}
\end{thm}
\juihsuan{maybe we should take $\form{\Gamma}$ instead? (consider the proof
$\vlderivation{\vlin{\wk}{}{\sqn{\ttt, A}}{
                 \vlin{\ttt}{}{\sqn\ttt}{
                   \vlhy{}}}}$)}

\begin{proof}
  Note that the instances of $\wrD,\cD$ in $\Deri_2$ are deep, but inside sequent contexts.
  \todo{}
\end{proof}

@@

\begin{thm}
  Any unification net can be obtained via the translation $\sqntl\cdot$ given in Theorem
  \ref{thm:MLL1->Unet}.
\end{thm}

@@

\begin{proof}
Suppose that such $\vlor$ node does not exist.
\todo{}
\end{proof}

\begin{lemma}
Suppose that $\lambda$ is a $\FOMLL$ proof net which is connected and such that any of its
switching graphs is not connected. Then there exists a $\vlor$ node in
$\unifstr{\lambda}$ such that $\lambda$ is correct on the sequent $\Gamma'$
obtained from $\Gamma$ by replacing this $\vlor$ by a $\vlan$.
\end{lemma}

@@

\begin{proof}\label{lem:concomp_switching_fo}
Let $\lambda$ be a unification net on $\Gamma$. We proceed by induction on the
number of connected components of the unification structure $\unifstr{\lambda}$:
\begin{itemize}
  \item If there is only one connected component, we proceed by induction on the
number $k$ of connected components of any switching graph of
$\unifstr{\lambda}$. If $k = 1$, we obtain a proof $\Deri$ in $\FOMLL$ such that
$\sqntl{\Deri} = \lambda$ by applying \cite[Theorem~3]{hughes:unifn}. If $k > 1$,
using the Lemma \ref{lem:concomp_switching_fo}, we obtain a sequent $\Gamma'$ on
which $\lambda$ is correct by transforming a $\vlor$ node into a $\vlan$. By
induction hypothesis, there is a proof $\Deri'$ in $\FOMLL$ whose translation is
$\lambda$. By considering the $\vlan$ rule instance corresponding to the $\vlan$
node in $\Deri'$, we have:
$
\Deri' =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma'}}{
  \vlin{}{}{\vlvdots}{
    \vliin{\vlan}{}{\sqns{\Delta_1, A \vlan B, \Delta_2}}{
      \vltrl{\Deri_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Deri_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}
$.
We can thus obtain a proof $\Deri$ of $\Gamma$:
$
\Deri =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma}}{
  \vlin{}{}{\vlvdots}{
    \vlin{\vlor}{}{\sqns{\Delta_1, A \vlor B, \Delta_2}}{
    \vliin{\mix}{}{\sqns{\Delta_1, A, B, \Delta_2}}{
      \vltrl{\Deri_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Deri_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}}
$ such that $\sqntl{\Deri} = \lambda$.

>>>>

%All statements in this section are proved by rule permutation arguments.
\begin{thm}\label{thm:LK1-decompose}
  Let $\Gamma$ be a sequent. If\/ $\sqn\Gamma$ is provable in $\FOLK$ (as depicted on the left below) then there is a sequent
  $\Gamma'$ such that there is a derivation as shown on the right below:\vadjust{\vskip-2ex}
  \begin{equation*}
    \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\Gamma}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}
    \qquad
    \Longrightarrow
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{\Deri_2}{\sqns{\form{\Gamma}}}{
        \vltrl{}{\FOMLL}{\Deri_1}{\sqns{\Gamma'}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}}}
  \end{equation*}
\end{thm}
\juihsuan{maybe there is still some problem... (consider the proof

$\vlderivation{
   \vliin{\vlan}{}{\sqn{\ttt, a \vlan b, \cneg b}}{
   \vlin{\ax}{}{\sqn{b, \cneg b}}{
     \vlhy{}}}{
   \vlin{\wk}{}{\sqn{\ttt, a}}{
     \vlin{\ttt}{}{\sqn\ttt}{
       \vlhy{}}}}}
$, it doesn't seem possible to get rid of the $\wk$ rule using deep weakening)}
\begin{proof}
  Note that the instances of $\wrD,\cD$ in $\Deri_2$ are deep, but inside sequent contexts.
  Suppose that $\Gamma$ is provable in $\FOLK$, then so is $\form{\Gamma}$.
Consider a proof
$ \vlderivation{
        \vltrl{}{\FOLK}{\Deri}{\sqns{\form{\Gamma}}}{
          \vlhy{\quad}}{
          \vlhy{}}{
          \vlhy{}}} $
Note that the instances of $\ctr$ and $\wk$ in $\Deri$ can be transformed into
derivations using $\cD$, $\wrD$ and the rules from $\FOMLL$.

\begin{equation*}
\vlderivation{
  \vlin{\ctr}{}{\sqns{\Gamma, A}}{
    \vlhy{\sqns{\Gamma, A, A}}}}
\leadsto
\vlderivation{
  \vlin{\cD}{}{\sqns{\Gamma, A}}{
    \vlin{\vlor}{}{\sqns{\Gamma, A \vlor A}}{
      \vlhy{\sqns{\Gamma, A, A}}}}}
\end{equation*}

\todo{}
\end{proof}

@@

\begin{thm}\label{thm:sqntl}
  Any unification net can be obtained via the translation $\sqntl\cdot$ given in Theorem
  \ref{thm:MLL1->Unet}.
\end{thm}

To prove this theorem, we need some basic lemmas about connected components in
switching graphs of unification nets.

@@

\begin{proof}
Suppose that such a $\vlor$ node does not exist. Then it is clear that for any
two nodes, there exists a switching graph containing a path between them and
this path corresponds to an $AE$-path in \cite{retore:03}. By
\cite[Propostion~3]{retore:03}, $\lambda$ corresponds to a sequent proof that
does not use $\mix$, which implies the connectedness of the switching graphs of
$\lambda$. Contradiction.
\tocheck{}
\end{proof}

\begin{lemma}\label{lem:concomp_switching_fo}
Suppose that $\lambda$ is a $\FOMLL$ proof net which is connected and such that any of its
switching graphs is not connected. Then there exists a $\vlor$ node in
$\unifstr{\lambda}$ such that $\lambda$ is correct on the sequent $\Gamma'$
obtained from $\Gamma$ by replacing this $\vlor$ by a $\vlan$.
\end{lemma}

@@

We can now give the proof of Theorem \ref{thm:sqntl}.

\begin{proof}[Proof of Theorem~\ref{thm:sqntl}]
Let $\lambda$ be a unification net on $\Gamma$. We proceed by induction on the
number of connected components of the unification structure $\unifstr{\lambda}$:
\begin{itemize}
  \item If there is only one connected component, we proceed by induction on the
number $k$ of connected components of any switching graph of
$\unifstr{\lambda}$. If $k = 1$, we obtain a proof $\Deri$ in $\FOMLL$ such that
$\sqntl{\Deri} = \lambda$ by applying \cite[Theorem~3]{hughes:unifn}. If $k > 1$,
using the Lemma \ref{lem:concomp_switching_fo}, we obtain a sequent $\Gamma'$ on
which $\lambda$ is correct by transforming a $\vlor$ node into a $\vlan$. By
induction hypothesis, there is a proof $\Deri'$ in $\FOMLL$ whose translation is
$\lambda$. By considering the $\vlan$ rule instance corresponding to the $\vlan$
node in $\Deri'$, we have:
$
\Deri' =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma'}}{
  \vlin{}{}{\vlvdots}{
    \vliin{\vlan}{}{\sqns{\Delta_1, A \vlan B, \Delta_2}}{
      \vltrl{\Deri_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Deri_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}
$.
We can thus obtain a proof $\Deri$ of $\Gamma$:
$
\Deri =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma}}{
  \vlin{}{}{\vlvdots}{
    \vlin{\vlor}{}{\sqns{\Delta_1, A \vlor B, \Delta_2}}{
    \vliin{\mix}{}{\sqns{\Delta_1, A, B, \Delta_2}}{
      \vltrl{\Deri_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Deri_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}}
$ such that $\sqntl{\Deri} = \lambda$.
